---
title: "Evaluations of positive and negative stimuli using the Affective Misattribution Procedure (AMP) and self-reports"
subtitle: "Data processing"
author: "Template: Ian Hussey; content: Nadia Almeida Ferreira "
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_folding: hide
    highlight: haddock
    theme: flatly
    toc: yes
    toc_float: yes
---

```{r, include=FALSE}

knitr::opts_chunk$set(message = FALSE,
                      warning = FALSE)

```

# Dependencies

```{r}

library(dplyr)
library(tidyr)
library(readr)
library(janitor) # for clean_names()
library(stringr)

```

# Get data

```{r}

# # demographics
# data_demographics_raw_messy <- read_csv("../data/raw/data_demographics_raw_messy.csv", skip = 2) |>
#   janitor::clean_names()

data_demographics_raw <- read_csv("../data/raw/data_demographics_raw.csv") |>
  janitor::clean_names()

# self report measure
data_selfreport_raw <- read_csv("../data/raw/data_selfreport_raw.csv") |>
  janitor::clean_names()

# affect attribution procedure
data_amp_raw <- read_csv("../data/raw/data_amp_raw.csv") |>
  janitor::clean_names()

```

# Duplicate unique_ids

```{r}

# do we have duplicate data
data_demographics_raw |>
  count(subject) |>
  arrange(desc(n))

```

yes - n is different for some subjects

```{r}

dat_unique_id <- data_demographics_raw |>
  rename(unique_id = subject) |>
  count(unique_id) |>
  filter(n == 2) |>
  select(unique_id)
  
```

# Demographics

Clean the demographics data so that summaries of age and gender can be made in the analysis.

```{r}

# wrangle the age data out of the raw
dat_age <- data_demographics_raw |>
  # rename for clarity
  rename(unique_id = subject, item = trialcode) |>
  # select useful columns
  select(unique_id, item, response) |>
  # retain rows that we need
  filter(item == "age") |>
  rename(age = response) |>
  # remove non-numbers or decimals from age
  mutate(age = str_remove_all(age, "[^\\d.]"),  # retains only digits and decimal points
         age = na_if(age, ""),
         age = as.numeric(age)) |>
  select(unique_id, age)
  #select(-item) # alternative: negative selection

# wrangle the gender data out of the raw
dat_gender <- data_demographics_raw |>
  # rename for clarity
  rename(unique_id = subject, item = trialcode) |>
  # select useful columns
  select(unique_id, item, response) |>
  # retain rows that we need
  filter(item == "gender") |>
  rename(gender = response) |>
  # remove non-numbers or decimals from age
  mutate(gender = str_to_lower(gender),
         gender = str_remove_all(gender, "[\\d.]"),
         gender = na_if(gender, ""), 
         gender = case_when(gender == "woman" ~ "female",
                            gender == "yes" ~ NA_character_,
                            gender == "man" ~ "male",
                            gender == "girl" ~ "female",
                            gender == "dude" ~ "male",
                            gender == "non binary" ~ "non-binary",
                            TRUE ~ gender)) |>
  select(unique_id, gender)

```

## Combine

```{r}

# dat_unique_id
# dat_age
# dat_gender

```

# Exclusions / data quality

## AMP

### Completeness

Create an exclusion variable `exclude_amp_completeness` based on the AMP having the right number of trials.

```{r}

# basic tidying
data_amp_tidied <- data_amp_raw |>
  rename(unique_id = subject,
         block_type = blockcode,
         rt = latency) |>
  select(unique_id, block_type, rt) |>
  filter(block_type == "test")

# are there duplicates or incomplete data?
data_amp_tidied |>
  count(unique_id, name = "n_trials") |>
  count(n_trials, name = "n_participants") |>
  arrange(desc(n_participants))

data_complete_amp <- data_amp_tidied |>
  count(unique_id, name = "n_trials") |>
  filter(n_trials == 72) 

data_amp_tidied_complete <- data_amp_tidied |>
  semi_join(data_complete_amp, by = "unique_id")

# are there duplicates or incomplete data?
# sanity check should have one row
data_amp_tidied_complete |>
  count(unique_id, name = "n_trials") |>
  count(n_trials, name = "n_participants") |>
  arrange(desc(n_participants))
  
# rm(data_complete_amp) # remove from environment

```

## Fast trials

Create an exclusion variable `exclude_amp_performance` based on AMP performance: "exclude" if more than 10% of the test block trials trials are < 100ms, or "include" if not.

```{r}

data_amp_performance_criteria <- data_amp_tidied_complete |>
  mutate(fast_trial = ifelse(rt < 100, 1, 0)) |>
  # for each participant...
  group_by(unique_id) |>
  # ... summarize the percent of fast trials
  summarise(percent_fast_trials = mean(fast_trial)*100) |>
  # flag unique_ids that have too many fast trials
  mutate(exclude_amp_performance = ifelse(percent_fast_trials > 10, "exclude", "include"))

```

## Self-reports

### Completeness

Create an exclusion variable `exclude_amp_completeness` based on the AMP having the right number of trials.

```{r}

#colnames(data_selfreport_raw)

# basic tidying
data_selfreport_tidied <- data_selfreport_raw |>
  rename(unique_id = subject,
         item_type = trialcode) |>
  select(unique_id, item_type, response) |>
  filter(item_type != "instructions")

```

TODO - 14/10

- replace all non 1-7 numbers in `response` with NA
- count number of unique_ids with correct/wrong number of responses
- Exclude them
- check your exclusions worked.
- use: mutate, group_by, summarize, _join

```{r}
#replace all non 1-7 numbers in `response` with NA

selfreport_response_tidied <- data_selfreport_tidied |>
  mutate(response = ifelse(response %in% 1:7, response, NA)) |> # put c("1", "2" ...) if all replaced with NA
  mutate(reponse = as.numeric(response)) #cuz response still a character and won't give you a mean for example

#count number of unique_ids with correct/wrong number of responses
selfreport_unique_ids_resultscount <- selfreport_response_tidied |>
  group_by(response) |>
  summarise(unique_count = n_distinct(unique_id)) # WRONG!!!!!
#Exclude them
selfreport_excluded_ids<- selfreport_response_tidied |>
  filter(is.na(response)) |>
  select(unique_id)

selfreport_excluded <- selfreport_response_tidied |>
  anti_join(selfreport_excluded_ids, by = "unique_id")

#check your exclusions worked
check_exclusions <- selfreport_excluded |>
  group_by(response)|>
  summarise(unique_ids_left= n_distinct(unique_id))




### Ian's version to negatively get rid of wrong answer
temp <- data_selfreport_tidied |>
  mutate(response = ifelse(response %in% 1:7, response, NA),
         response = as.numeric(response)) #could put all in one line by putting response = as.numeric(ifelse(response.....))


class(temp$response)

mean(temp$response, na.rm= TRUE)

# count how many trials each unique_id has 

temp |>
  count(unique_id, name = "n_trials") |>
  count(n_trials, name = "n_participants")

# count how many trials each unique_id has 

temp |>
  group_by(unique_id)|>
  summarize(n_trials = n()) |>
  group_by(n_trials) |>
  summarise(n_participans = n())

# exclude them (the ones with not 3 trials )
# temp2 <-temp|>
#   group_by(unique_id) |>
#   mutate(n_trials = n()) |>
#   filter(n_trials ==3)

unique_id_with_3_items <-temp |>
  count(unique_id) |>
  filter(n==3)


data_selfreport_complete <- 
  # inner_join(data_selfreport_tidied,
  #            unique_id_with_3_items,
  #            by = "unique_id") #keeps all collums
  semi_join(data_selfreport_tidied,
        unique_id_with_3_items,
        by = "unique_id") #does not show the unique id n column 
```

# Score measures

## Self-reports

Create a mean score for each participant of the three self-report items.

```{r}

data_selfreport_mean_score <- data_selfreport_complete|>
  mutate(response = as.numeric(response)) |>
  group_by(unique_id) |>
  summarize(mean_response = mean(response))



```

## Affect Misattribution Procedure

Create an overall AMP bias score. Score each trial as having been evaluated as congruent with the prime or not. eg When trialcode == "prime_positive" & evaluative_response == 1, then record as 1, etc. AMP score is the proportion of congruent responses, using the test blocks only.

```{r}
#calculate mean and sd of reaction times for each perticipant
data_amp_score_congruence <- data_amp_tidied_complete |>
  group_by(unique_id) 
  
data_amp_tidied_complete |>
  count(unique_id) |>
  count(n)

```

# Combine 

Combine data frames together to create one wide-format data frame where each row represents a participant.

Flag all duplicates for exclusion.

```{r}

# data_processed_before_exclusions 

```

# Define master exclusion variable

```{r}

#data_processed <- data_processed_before_exclusions

```

# Write to disk

```{r}

# # in case this dir doesn't exist, create it
# dir.create("../data/processed/")
# 
# # save data to disk in that dir
# write_csv(data_processed, "../data/processed/data_processed.csv")

```

# Session info

```{r}

sessionInfo()

```


